### 2009-03-20 doc/storage_system/arc-storage-manual.tex
manual about adding more storage elements
TODO: manual about adding more Librarians and Bartenderds
TODO: manual about changing the CentralAHash to ReplicatedAHash

\chapter{Administrator's manual} % (fold)
\label{cha:administrator_s_manual}

\section{Quick start guide} % (fold)
\label{sec:quick_start_guide}

@@

% section quick_start_guide (end)

% \section{Installation} % (fold)
% \label{sec:installation}
% 
% % section installation (end)
% 
% \section{Configuration} % (fold)
% \label{sec:configuration}
% 
% % section configuration (end)
% 
% \section{Deployment scenarios} % (fold)
% \label{sec:deployment_scenarios}
% 
% % section deployment_scenarios (end)

% chapter administrator_s_manual (end)

\chapter{User's manual} % (fold)
\label{cha:user_s_manual}

@@

\begin{verbatim}
$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \\
arc_storage_cli list /
- Calling the Bartender's list method...
- done in 0.31 seconds.
'/': denied

$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \\
arc_storage_cli put /tmp/orange /pennys-orange
- The size of the file is 7 bytes
- Calculating md5 checksum...
- done in 0.0001 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
- Calling the Bartender's putFile method...
- done in 0.81 seconds.
/pennys-orange: failed to add child to parent

@@

\begin{verbatim}
$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \\
arc_storage_cli put /tmp/orange /pennys-orange
- The size of the file is 7 bytes
- Calculating md5 checksum...
- done in 0.0001 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
- Calling the Bartender's putFile method...
- done in 1.18 seconds.
- Got transfer URL: http://localhost:60001/hopi/4145891d-0e4c-b6bd-e4d6-dc80554b9d35
- Uploading from '/tmp/orange'
    to 'http://localhost:60001/hopi/4145891d-0e4c-b6bd-e4d6-dc80554b9d35' with http...
    1 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
- done in 1.0071 seconds.
'/tmp/orange' (7 bytes) uploaded as '/pennys-orange'.
\end{verbatim}

@@

\begin{verbatim}
$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \\
arc_storage_cli pol /pennys-orange change /DC=eu/DC=KnowARC/O=knowarc/CN=billy +read
- Calling the Bartender's stat method...
- done in 0.41 seconds.
old action list for the user /DC=eu/DC=KnowARC/O=knowarc/CN=billy was ''
the new action list of user /DC=eu/DC=KnowARC/O=knowarc/CN=billy for LN /pennys-orange
    will be '+read'
- Calling the Bartender's modify method...
- done in 0.59 seconds.
set.

@@

We can see that the `neededReplica' metadata says 2, but if we deployed the system in a centralized way for testing purposes, then we only have one storage element (which is a Shepherd and a Hopi service together), so the file can only have one replica, that's why there is only one locations.

@@

% \section{Overview} % (fold)
% \label{sec:overview}
% 
% % section overview (end)
% 
% \section{Prototype command line tool} % (fold)
% \label{sec:prototype_command_line_tool}
% 
% % section prototype_command_line_tool (end)
% 
% \section{Grid jobs} % (fold)
% \label{sec:grid_jobs}
% 
% % section grid_jobs (end)
% 
% \section{Accessing the system with FUSE} % (fold)
% \label{sec:accessing_the_system_with_fuse}
% 
% % section accessing_the_system_with_fuse (end)

% chapter user_s_manual (end)

% \bibliography{grid}
\end{document}

>>>>

\chapter{Quick start guide} % (fold)

\section{Installing a centralized system} % (fold)

@@

\section{Using the prototype client tool} % (fold)

@@

\begin{verbatim}
$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \
arc_storage_cli list /
- Calling the Bartender's list method...
- done in 0.31 seconds.
'/': denied

$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \
arc_storage_cli put /tmp/orange /pennys-orange
- The size of the file is 7 bytes
- Calculating md5 checksum...
- done in 0.0001 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
- Calling the Bartender's putFile method...
- done in 0.81 seconds.
/pennys-orange: failed to add child to parent

@@

\begin{verbatim}
$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \
arc_storage_cli put /tmp/orange /pennys-orange
- The size of the file is 7 bytes
- Calculating md5 checksum...
- done in 0.0001 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
- Calling the Bartender's putFile method...
- done in 1.18 seconds.
- Got transfer URL: http://localhost:60001/hopi/4145891d-0e4c-b6bd-e4d6-dc80554b9d35
- Uploading from '/tmp/orange'
    to 'http://localhost:60001/hopi/4145891d-0e4c-b6bd-e4d6-dc80554b9d35' with http...
    1 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
- done in 1.0071 seconds.
'/tmp/orange' (7 bytes) uploaded as '/pennys-orange'.
\end{verbatim}

@@

\begin{verbatim}
$ ARC_CERT_FILE=usercert-penny.pem ARC_KEY_FILE=userkey-penny.pem \
arc_storage_cli pol /pennys-orange change /DC=eu/DC=KnowARC/O=knowarc/CN=billy +read
- Calling the Bartender's stat method...
- done in 0.41 seconds.
old action list for the user /DC=eu/DC=KnowARC/O=knowarc/CN=billy was ''
the new action list of user /DC=eu/DC=KnowARC/O=knowarc/CN=billy for LN /pennys-orange
    will be '+read'
- Calling the Bartender's modify method...
- done in 0.59 seconds.
set.

@@

We can see that the `neededReplica' metadata says 2, but if we deployed the system in a centralized way for testing purposes, then we only have one storage element (which is a Shepherd service and a Hopi service together), so the file can only have one replica, that's why there is only one locations.

@@

\section{Adding more storage elements}

Let's add a new storage element to the system. We need to deploy a Shepherd service and a storage element service, which this time will be a `byteio' service. The `byteio' service implements a small part of the ByteIO interface, which means that it just put the entire file data into the SOAP message, so it is only suitable for small files.

But first modify the configuration file of our original server. Currently all the transfer URLs which were generated when we upload and download files contained the hostname `localhost'. This is not OK if we want to access it from an other machine, we should provide transfer URLs which contains the external hostname of the machine. So lets change this line in \verb!/etc/arc/storage_service.xml!:

\begin{verbatim}
    <TURLPrefix>http://localhost:60001/hopi/</TURLPrefix>
\end{verbatim}

to the real hostname of the machine. But this is not enough. Currently the Shepherd services uses their URLs as IDs, and the other services use these IDs to connect to a Shepherd service, so we need to change to ID of the Shepherd service to a real external URL as well. And for the sake of clarity let's change all `localhost' entries in the config file to our real hostname. In our case the name of this machine is just `storage' so the Shepherd section should look like this:

\begin{verbatim}
<Service name="pythonservice" id="shepherd">
    <ClassName>storage.shepherd.shepherd.ShepherdService</ClassName>
    <ServiceID>https://storage:60000/Shepherd</ServiceID>
    <CheckPeriod>20</CheckPeriod>
    <MinCheckInterval>0.1</MinCheckInterval>
    <StoreClass>arcom.store.cachedpicklestore.CachedPickleStore</StoreClass>
    <StoreCfg><DataDir>/var/spool/arc/shepherd_data</DataDir></StoreCfg>
    <BackendClass>storage.shepherd.hardlinkingbackend.HopiBackend</BackendClass>
    <BackendCfg>
        <DataDir>/var/spool/arc/shepherd_store</DataDir>
        <TransferDir>/var/spool/arc/shepherd_transfer</TransferDir>
        <TURLPrefix>http://storage:60001/hopi/</TURLPrefix>
    </BackendCfg>
    <LibrarianURL>https://storage:60000/Librarian</LibrarianURL>
    <BartenderURL>https://storage:60000/Bartender</BartenderURL>
    <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
</Service>
\end{verbatim}

Let's check first if our centralized deployment is still working, stop the \verb!arched! daemon remove the data directories to make it a clean start, and restart it with the modified config.

\begin{verbatim}
$ sudo rm -r /var/spool/arc/*
$ sudo /usr/local/sbin/arched -c /etc/arc/storage_service.xml -f    
\end{verbatim}

Let's see if we can still use the client:

\begin{verbatim}
$ arc_storage_cli list /
- Calling the Bartender's list method...
- done in 0.31 seconds.
'/': not found

$ arc_storage_cli mkdir /
- Calling the Bartender's makeCollection method...
- done in 0.63 seconds.
Creating collection '/': done

$ arc_storage_cli list /
- Calling the Bartender's list method...
- done in 0.50 seconds.
'/': collection
    empty.
\end{verbatim}

Let's upload a file and ask for two replicas:

\begin{verbatim}
$ ARC_NEEDED_REPLICAS=2 arc_storage_cli put orange /
- The size of the file is 7 bytes
- Calculating md5 checksum...
- done in 0.0005 seconds, md5 checksum is 6aefd2842be62cd470709b27aedc7db7
- Calling the Bartender's putFile method...
- done in 1.10 seconds.
- Got transfer URL: http://storage:60001/hopi/139dc57d-8c12-c188-a4a6-973449553811
- Uploading from 'orange'
    to 'http://storage:60001/hopi/139dc57d-8c12-c188-a4a6-973449553811' with http...
    1 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
- done in 1.0090 seconds.
'orange' (7 bytes) uploaded as '/orange'.

$ arc_storage_cli stat /orange
- Calling the Bartender's stat method...
- done in 0.41 seconds.
'/orange': found
  states
    checksumType: md5
    neededReplicas: 2
    size: 7
    checksum: 6aefd2842be62cd470709b27aedc7db7
  entry
    owner: /DC=eu/DC=KnowARC/O=knowarc/CN=billy
    GUID: 3fdad88a-944f-79c5-4f23-c5b3d3bc981e
    type: file
  parents
    0/orange: parent
  locations
    https://storage:60000/Shepherd 0b46f53c-419d-ce4e-181a-ec95d16dcc95: alive
  timestamps
    created: 1237519441.96
\end{verbatim}

Currently it has only one replica, because we still don't have more than one storage elements. While \verb!arched! keeps running, let's create a new config for the other machine, which is called \verb!storage2!:

\begin{verbatim}
<?xml version="1.0"?>
<ArcConfig 
  xmlns="http://www.nordugrid.org/schemas/ArcConfig/2007"
  xmlns:tcp="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"
>
    <Server>
        <Pidfile>/var/run/arched2.pid</Pidfile>
        <Logger level="ERROR">/var/log/arched2.log</Logger>
    </Server>
    <ModuleManager>
        <Path>/usr/local/lib/arc/</Path>
    </ModuleManager>
    <Plugins><Name>mcctcp</Name></Plugins>
    <Plugins><Name>mcctls</Name></Plugins>
    <Plugins><Name>mcchttp</Name></Plugins>
    <Plugins><Name>mccsoap</Name></Plugins>
    <Chain>
        <Component name="tcp.service" id="tcp">
            <next id="tls"/> 
            <tcp:Listen>
                <tcp:Port>60006</tcp:Port>
                <tcp:Version>4</tcp:Version>
            </tcp:Listen>
        </Component>
        <Component name="tls.service" id="tls">
            <next id="http"/>
            <KeyPath>/etc/grid-security/hostkey.pem</KeyPath>
            <CertificatePath>/etc/grid-security/hostcert.pem</CertificatePath>
            <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
        </Component>
        <Component name="http.service" id="http">
            <next id="soap">POST</next>
        </Component>
        <Component name="soap.service" id="soap">
            <next id="plexer"/>
        </Component>
        <Plexer name="plexer.service" id="plexer">
            <next id="shepherd">^/Shepherd$</next>
            <next id="byteio">^/byteio/</next>
        </Plexer>
        <Service name="pythonservice" id="shepherd">
            <ClassName>storage.shepherd.shepherd.ShepherdService</ClassName>
            <ServiceID>https://storage2:60006/Shepherd</ServiceID>
            <CheckPeriod>20</CheckPeriod>
            <MinCheckInterval>0.1</MinCheckInterval>
            <StoreClass>arcom.store.cachedpicklestore.CachedPickleStore</StoreClass>
            <StoreCfg>
                <DataDir>/var/spool/arc/shepherd_store2</DataDir>
            </StoreCfg>
            <BackendClass>storage.shepherd.byteio.ByteIOBackend</BackendClass>
            <BackendCfg>
                <DataDir>/var/spool/arc/shepherd_data2</DataDir>
                <TransferDir>/var/spool/arc/shepherd_transfer2</TransferDir>
                <TURLPrefix>https://storage2:60006/byteio/</TURLPrefix>
            </BackendCfg>
            <LibrarianURL>https://storage:60000/Librarian</LibrarianURL>
            <BartenderURL>https://storage:60000/Bartender</BartenderURL>
            <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml"/>
        </Service>
        <Service name="pythonservice" id="byteio">
            <ClassName>storage.shepherd.byteio.ByteIOService</ClassName>
            <TransferDir>/var/spool/arc/shepherd_transfer2</TransferDir>
            <NotifyURL>https://storage2:60006/Shepherd</NotifyURL>
            <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml"/>
        </Service>
    </Chain>
</ArcConfig>    
\end{verbatim}

This configuration contains only two services: a Shepherd, and byteio service. The byteio service is just a very simple storage element service which moves files within SOAP messages. The Shepherd service needs to have a backend for all supported storage element services, so it has a ByteIOBackend configured. This Shepherd connects to our original server's Bartender and Librarian services. (See the technical documentation for details about the services.)

While the original (\verb!storage!) server still runs the services, let's start this new server:

\begin{verbatim}
$ sudo /usr/local/sbin/arched -c /etc/arc/storage_service_2.xml -f
\end{verbatim}

And after a minute or two, let's check our file:

\begin{verbatim}
$ arc_storage_cli stat /orange
- Calling the Bartender's stat method...
- done in 0.41 seconds.
'/orange': found
  states
    checksumType: md5
    neededReplicas: 2
    size: 7
    checksum: 6aefd2842be62cd470709b27aedc7db7
  entry
    owner: /DC=eu/DC=KnowARC/O=knowarc/CN=billy
    GUID: 24b442eb-ab1a-680e-1f79-4c20479d3ea9
    type: file
  parents
    0/orange: parent
  locations
    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: alive
  timestamps
    created: 1237519828.4
\end{verbatim}

It indeed has two replicas now, one on each server.

Let's download it multiple times:

\begin{verbatim}
$ arc_storage_cli get /orange /tmp/orange1
- Calling the Bartender's getFile method...
- done in 0.50 seconds.
- Got transfer URL: http://storage:60001/hopi/de0d951f-7379-1ce4-0813-48a384c11982
- Downloading from 'http://storage:60001/hopi/de0d951f-7379-1ce4-0813-48a384c11982'
    to '/tmp/orange1' with http...
    0 s:        0.0 kB       0.0 kB/s       0.0 kB/s    . . .       
- done in 1.0060 seconds.
'/orange' (7 bytes) downloaded as '/tmp/orange1'.

$ arc_storage_cli get /orange /tmp/orange2
- Calling the Bartender's getFile method...
- done in 0.51 seconds.
- Got transfer URL: https://storage2:60006/byteio/52b6bffa-4812-03a5-93a1-834e38077ac5
- Downloading from 'https://storage2:60006/byteio/52b6bffa-4812-03a5-93a1-834e38077ac5'
    to '/tmp/orange2' with byteio...
- done in 0.2822 seconds.
'/orange' (7 bytes) downloaded as '/tmp/orange2'.
\end{verbatim}

First it was downloaded from the \verb!storage! server via HTTP, second time it was downloaded from the \verb!storage2! server via ByteIO.

Let's create a third storage element on a machine called \verb!storage3! this time using a Hopi service as storage element service like on our main \verb!storage! server:

\begin{verbatim}
zsombor@storage:~$ cat /etc/arc/storage_service_3.xml 
<?xml version="1.0"?>
<ArcConfig 
  xmlns="http://www.nordugrid.org/schemas/ArcConfig/2007"
  xmlns:tcp="http://www.nordugrid.org/schemas/ArcMCCTCP/2007"
>
    <Server>
        <Pidfile>/var/run/arched3.pid</Pidfile>
        <Logger level="ERROR">/var/log/arched3.log</Logger>
    </Server>
    <ModuleManager>
        <Path>/usr/local/lib/arc/</Path>
    </ModuleManager>
    <Plugins><Name>mcctcp</Name></Plugins>
    <Plugins><Name>mcctls</Name></Plugins>
    <Plugins><Name>mcchttp</Name></Plugins>
    <Plugins><Name>mccsoap</Name></Plugins>
    <Chain>
        <Component name="tcp.service" id="tcp">
            <next id="tls"/> 
            <tcp:Listen>
		<tcp:Port>60010</tcp:Port>
		<tcp:Version>4</tcp:Version>
	    </tcp:Listen>
        </Component>
        <Component name="tls.service" id="tls">
            <next id="http"/>
            <KeyPath>/etc/grid-security/hostkey.pem</KeyPath>
            <CertificatePath>/etc/grid-security/hostcert.pem</CertificatePath>
            <CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
        </Component>
        <Component name="http.service" id="http">
            <next id="soap">POST</next>
        </Component>
        <Component name="soap.service" id="soap">
            <next id="plexer"/>
        </Component>
        <Plexer name="plexer.service" id="plexer">
            <next id="shepherd">^/Shepherd$</next>
        </Plexer>
        <Service name="pythonservice" id="shepherd">
            <ClassName>storage.shepherd.shepherd.ShepherdService</ClassName>
            <ServiceID>https://storage3:60010/Shepherd</ServiceID>
            <CheckPeriod>20</CheckPeriod>
            <MinCheckInterval>0.1</MinCheckInterval>
            <StoreClass>arcom.store.cachedpicklestore.CachedPickleStore</StoreClass>
            <StoreCfg><DataDir>/var/spool/arc/shepherd_data3</DataDir></StoreCfg>
            <BackendClass>storage.shepherd.hardlinkingbackend.HopiBackend</BackendClass>
            <BackendCfg>
                <DataDir>/var/spool/arc/shepherd_store3</DataDir>
                <TransferDir>/var/spool/arc/shepherd_transfer3</TransferDir>
                <TURLPrefix>http://storage3:60011/hopi/</TURLPrefix>
            </BackendCfg>
            <LibrarianURL>https://storage:60000/Librarian</LibrarianURL>
            <BartenderURL>https://storage:60000/Bartender</BartenderURL>
            <ClientSSLConfig FromFile="/etc/arc/clientsslconfig.xml" />
        </Service>
    </Chain>
    <Chain>
        <Component name="tcp.service" id="tcp2">
            <next id="http2"/> 
            <tcp:Listen>
		<tcp:Port>60011</tcp:Port>
		<tcp:Version>4</tcp:Version>
	    </tcp:Listen>
        </Component>
        <Component name="http.service" id="http2">
            <next id="plexer2">GET</next>
            <next id="plexer2">PUT</next>
        </Component>
        <Plexer name="plexer.service" id="plexer2">
            <next id="hopi">^/hopi/</next>
        </Plexer>
        <Service name="hopi" id="hopi">
            <DocumentRoot>/var/spool/arc/shepherd_transfer3</DocumentRoot>
            <SlaveMode>1</SlaveMode>
        </Service>
    </Chain>
</ArcConfig>    
\end{verbatim}

This configuration has two separate ports, one with TLS security and a Shepherd service, and the other without TLS security and with the Hopi service. Let's run this server as well:

\begin{verbatim}
$ sudo /usr/local/sbin/arched -c /etc/arc/storage_service_3.xml -f
\end{verbatim}

Now when checking our file again there should be no change at all:

\begin{verbatim}
$ arc_storage_cli stat /orange
  locations
    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: alive
\end{verbatim}

Let's stop our server on \verb!storage2!, and watch what happens of our file:

\begin{verbatim}
$ arc_storage_cli stat /orange
  locations
    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: offline

$ arc_storage_cli stat /orange
  locations
    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: offline
    https://storage3:60010/Shepherd 421aa939-81fa-5cd9-7c2f-2d5f99a98be1: creating

$ arc_storage_cli stat /orange
  locations
    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: offline
    https://storage3:60010/Shepherd 421aa939-81fa-5cd9-7c2f-2d5f99a98be1: alive
\end{verbatim}

A new replica have been created on the third server to maintain the needed number of 2.

Let's start the services on \verb!storage2! again, and let's see what is happening with our file:

\begin{verbatim}
$ arc_storage_cli stat /orange
  locations
    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: alive
    https://storage3:60010/Shepherd 421aa939-81fa-5cd9-7c2f-2d5f99a98be1: alive

$ arc_storage_cli stat /orange
  locations
    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
    https://storage2:60006/Shepherd c1334c09-5d3d-983b-2081-87fe60d4764a: thirdwheel
    https://storage3:60010/Shepherd 421aa939-81fa-5cd9-7c2f-2d5f99a98be1: alive

$ arc_storage_cli stat /orange
  locations
    https://storage:60000/Shepherd 315d59d3-50cd-1a10-02e6-8e6a528473c3: alive
    https://storage3:60010/Shepherd 421aa939-81fa-5cd9-7c2f-2d5f99a98be1: alive
\end{verbatim}

One of the replicas have been disappeared to maintain the needed number of 2.

\end{document}
